Robust Statistical Stopping Criteria for Automated Screening in Systematic Reviews

### 2020-06-13 manuscript/main.tex
resubmission

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass[linenumbers]{bmcart}

@@

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{url}
\usepackage{subcaption}
\usepackage{booktabs}
%\def\includegraphic{}
%\def\includegraphics{}

@@

	Evidence synthesis technology is a rapidly emerging field that promises to change the practice of evidence synthesis work \cite{Westgate2018}.
	Interventions have been proposed at various points in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
	A major strand of the literature works on screening: the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
	This is a time consuming and repetitive task, and in a research environment with constrained resources and increasing amounts of literature, this may limit the scope of the evidence synthesis projects undertaken.
	Several papers have developed Active Learning (AL) approaches \cite{miwa2014, Wallace2010a, Wallace2010, Jonnalagadda2013, Przybya2018} to reduce the time required to screen documents. This paper sets out how current approaches are  unsuitable in practice, and outlines and evaluates a small modification that would make AL systems ready for live reviews.

	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers \cite{Settles2009}.
	The algorithm chooses which studies will next be screened by humans, often those which are likely to be relevant or about which the model is uncertain, in order to generate more labels to feed back to the machine. 
	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (described by recall: the number of relevant studies identified divided by the total number of relevant studies) - before having seen all the documents in the corpus. 
	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved. This represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.

@@

	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall. If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of 95\%''.

	In the remainder of the paper, we first discuss in detail the shortcomings of existing stopping criteria. Then, we introduce our new criterion based on a hypergeometric test. We evaluate our stopping criteria, and compare their performance with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.

@@

	However, the estimation of the BIR fails to correctly take into account sampling uncertainty \footnote{Although Shemilt et al. \cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choice. Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 451$ relevant results. To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results. See the notebook \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb} that accompanies this paper for a detailed discussion}. 
	This uncertainty is crucial, as errors can have severe consequences. If the estimated number of relevant documents is even one unit above the true value, then no work savings will be achieved. If the number of relevant documents is underestimated, then the recall achieved will be less than 100\%.

	The number of relevant documents drawn without replacement from a finite sample of documents follows the hypergeometric distribution. 
	Figure \ref{fig:bir_error} shows the distribution of the predicted number of documents after drawing 1,000 documents from a total of 20,000 documents, where 500 documents (2.5\%) are relevant. The left shaded portion of the graph shows all the cases where the recall will be less than 95\%. This occurs 35.91\% of the time. The right shaded portion of the graph shows the cases where the number of relevant documents is overestimated and no work savings could be made. This occurs 46.24\% of the time. In only 17.85\% of cases can work savings be achieved while still achieving a recall of at least 95\%.

	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size. Even in very large samples both types of error remain frequent, and the risk of saving no work at all remains close to 50\%.
	This shows how baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.

@@

Some studies give the example of heuristic stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low, as the probability of observing 0 relevant documents in a given sample (analogous to a set of consecutive irrelevant results) is inversely related to the number of relevant documents in the population.
We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; 
2) it lacks a formal definition; and 3) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.

Figure \ref{unseen-proportions} illustrates this third point. 
We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
In the top figure, machine learning (ML) has performed well, and 74\% of the screened documents were relevant. 
In the bottom figure,  ML has performed less well, and only 26\% of the screened documents were relevant.
In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. 
This also means that where ML has performed well (as in the top figure), a low proportion of relevant documents in those not yet checked is indicative of lower recall than where ML has performed less well.
Likewise, where the proportion of relevant documents in the whole corpus is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low. 
This shows us that even a perfect estimator of the relevance of unseen documents is insufficient on its own to provide sufficient information about when to stop screening.

@@

Finally, Howard et al. \cite{Howard2020} present a method to estimate recall based on the the number of irrelevant documents $D$ observed in a list of documents since the $\delta th$ previous relevant document. 
They reason that this should follow the negative binomial distribution based on the proportion of remaining relevant documents $p$, and use this information to estimate $\hat{p}$, and with this, the total number of relevant articles and the estimated recall. 
However, since screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial. Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$. Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally. Lastly, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours).

These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance. In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.

@@

	In our screening setup, we start off with $N_T$ documents that are potentially relevant. $\rho_{tot}$ of these documents are actually relevant, but we don't know this value \textit{a priori}. As we screen relevant documents we include them, so $\rho_{seen}$ represents the number of relevant documents screened, and $\tau$ recall is given by

	\begin{equation}
		\tau = \frac{\rho_{seen}}{\rho_{tot}}
	\end{equation}

@@

	At the start of the sample, $N_{AL}$ is the number of documents seen during the active learning process, and $N_s$ is the number of documents remaining, so that

	\begin{equation}
	N_s = N_T - N_{AL}
	\end{equation}

@@

	\begin{equation}
	K = \rho_{tot} - \rho_{AL}
	\end{equation}

	With each draw, $n$ is the number of documents drawn and $k$ is the number of relevant documents drawn. The number of relevant documents seen is updated by adding the number of relevant documents seen since sampling began to the number of relevant documents seen during active learning.

	\begin{equation}
	\rho_{seen} = \rho_{ML} + k
	\end{equation}

@@

	\begin{equation}
	H_0 : \tau < \tau_{tar}
	\end{equation}

@@

	\begin{equation}
	X \sim Hypergeometric(N, K, n)
	\end{equation}

	We introduce a hypothetical value for $K$, which we call $K_0$. This represents the minimum number of relevant documents remaining at the start of sampling compatible with our null hypothesis that recall is below our target.

	\begin{equation}
	K_0 = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
	\end{equation}

	Our null hypothesis can thereby reformulated: the true number of relevant documents in the sample is greater than our hypothetical value.

@@

	We test this by calculating the probability of observing $k$ or fewer relevant from the hypergeometric distribution given by $K_{tar}$, using the cumulative distribution function.

	\begin{equation}
	p = P(X \leq k)$, where $X \sim Hypergeometric(N_s,K_{tar},n)
	\end{equation}

	Because $P(X \leq k)$, is strictly decreasing for values below $K_{tar}$, this gives the maximum probability of observing $k$ for all values of K compatible with our null hypothesis.

	If the maximum probability of obtaining our observed results given our null hypothesis is below our $1-\alpha$, then we can reject our null hypothesis and stop screening. We can report the likelihood that we achieve a recall below our target as being less than $1-\alpha$.

	\subsubsection*{Nonrandom sampling}

	In order to decide when to begin a random sample, we employ nonrandom sampling, where we treat previously screened documents as random samples. The distribution of relevant documents among previously screened documents is clearly not random, as documents predicted to be relevant are prioritised. It is reasonable to assume, though, that the density of relevant documents is greater among previously screened documents than among remaining unseen documents. This would make the following estimates conservative.

	After reviewing each document, $N_{seen}$ documents have been screened, and $N_{unseen}$ documents are yet to be seen. 
	We treat $i = 1 \dots N_{seen}$ of the previously screened documents as separate random samples, and calculate $p$, using the method above, for each sample. 
	We take the minimum across all samples $p_{min}$ as the lowest probability that our null hypothesis is correct. 
	If $p_{min}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling. 
	This method is an imprecise heuristic used to decide when to start random sampling. 
	However, we also test its performance as a separate stopping criterion.
	We calculate $p_{min}$ for the remaining documents as if we had not switched to random sampling and record the recall and work saved when $p_{min} < 1 - \alpha$. 
	We present these in the results below as the nonrandom sampling criterion.

@@

    \begin{figure*}
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_hyper.pdf}
		\caption[]%
		{{\small Hypergeometric sampling \\}}    
		\label{fig:hyper}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_nrs.pdf}
		\caption[]%
		{{\footnotesize Pseudorandom hypergeometric sampling}}    
		\label{fig:nrs}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{2_figs_jointplot_bir.pdf}
		\caption[Network2]%Example-Image
		{{\small Baseline Sampling}}    
		\label{fig:bir}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}  
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_ih_50.pdf}
		\caption[]%
		{{\small 50 consecutive irrelevant results}}    
		\label{fig:ih_50}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_ih_200.pdf}
		\caption[]%
		{{\small 200 consecutive irrelevant results \\}}    
		\label{fig:ih_200}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_pf.pdf}
		\caption[]%
		{{\footnotesize \textit{a priori} knowledge}}   
		\label{fig:pf}
	\end{subfigure}

@@

\begin{figure*}
	\centering
		\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_hyper.pdf}
		\caption[]%
		{{\small Hypergeometric sampling \\}}    
		\label{fig:hyper_ab}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_nrs.pdf}
		\caption[]%
		{{\footnotesize Pseudorandom hypergeometric sampling}}    
		\label{fig:nrs_ab}
	\end{subfigure}
	\hfill
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_bir.pdf}
		\caption[Network2]%Example-Image
		{{\small Baseline Sampling}}    
		\label{fig:bir_ab}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}  
		\centering 
		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_ih_50.pdf}
		\caption[]%
		{{\small 50 consecutive irrelevant results}}    
		\label{fig:ih_50_ab}
	\end{subfigure}
	\vskip\baselineskip

@@

	Both the random sampling and the pseudorandom sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. In fact, the pseudorandom sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases. In theory, the pseudorandom sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random. Based on our experiments, this assumption seems reasonable, and accounts for the higher recall. Because the pseudorandom sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data. This accounts for the higher work savings.

@@

	\begin{figure}
		\includegraphics[width=0.9\linewidth]{2_figs_wss_nrs.pdf}
		\caption{Work saved for the pseudorandom sampling method in each dataset. Labels show the number of relevant documents and the total number of documents. The datasets are presented in order of the number of documents. The whiskers represent the 5th and 95th percentiles. The grey line shows work savings of 5\%. }
		\label{wss}
	\end{figure}

@@

	To help explain the different work savings that were observed in our experiments, we show the distribution of work savings from our pseudorandom criterion for each dataset in figure \ref{wss}. In general, higher work savings are possible when the total number of documents is larger. However, in datasets with a low proportion of relevant documents, many documents need to be screened to achieve a high confidence that there are only few relevant documents remaining in the unseen ones. Therefore, smaller work savings are possible.

@@

>>>>

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

@@

\usepackage{todonotes}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{url}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{booktabs}
%\def\includegraphic{}
%\def\includegraphics{}

@@

	Evidence synthesis technology is a rapidly emerging field that promises to change the practice of evidence synthesis work \cite{Westgate2018}.
	Interventions have been proposed at various points in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
	A major strand of the literature works on screening: the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
	This is a time consuming and repetitive task, and in a research environment with constrained resources and increasing amounts of literature, this may limit the scope of the evidence synthesis projects undertaken.
	Several papers have developed Active Learning (AL) approaches \cite{miwa2014, Wallace2010a, Wallace2010, Jonnalagadda2013, Przybya2018} to reduce the time required to screen documents. This paper sets out how current approaches are  unreliable in practice, and outlines and evaluates modifications that would make AL systems ready for live reviews.

	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers \cite{Settles2009}.
	The algorithm chooses which studies will next be screened by humans, often those which are likely to be relevant or about which the model is uncertain, in order to generate more labels to feed back to the machine. 
	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies -- or a given proportion of relevant studies (described by recall: the number of relevant studies identified divided by the total number of relevant studies) -- before having seen all the documents in the corpus. 
	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved. This represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.

@@

	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall. If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of more than 95\%''.

	In the remainder of the paper, we first discuss in detail the shortcomings of existing stopping criteria. Then, we introduce our new criteria based on a hypergeometric test. We evaluate our stopping criteria, and compare their performance with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.

@@

	However, the estimation of the BIR fails to correctly take into account sampling uncertainty \footnote{Although Shemilt et al. \cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choice. Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 451$ relevant results. To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results. See the notebook \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb} that accompanies this paper for a detailed discussion.}. 
	This uncertainty is crucial, as errors can have severe consequences. Let us assume that users will stop screening when they have identified 95\% of the relevant number of documents. If the estimated number of relevant documents is more than the true number of relevant documents divided by 0.95, then the users will never see 95\% of the estimated number.  This means that they will keep screening until they have seen all documents, and no work savings will be achieved. Conversely, if the number of relevant documents is underestimated by even a single unit, then the recall achieved will be lower than the target.

	The number of relevant documents drawn without replacement from a finite sample of documents follows the hypergeometric distribution. 
	Figure \ref{fig:bir_error} shows the distribution of the predicted number of documents after drawing 1,000 documents from a total of 20,000 documents, where 500 documents (2.5\%) are relevant. The left shaded portion of the graph shows all the cases where the recall will be less than 95\%. This occurs 48\% of the time. The right shaded portion of the graph shows the cases where the number of relevant documents is overestimated so much that no work savings could be made to achieve a target recall of 95\%. This occurs 29\% of the time. In only 23\% of cases can work savings be achieved while still achieving a recall of at least 95\%.

	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size. Even in very large samples both types of error remain frequent.
	This shows how baseline estimation inevitably offers poor reliability, either in terms of recall or in work saved.

@@

	Some studies give the example of heuristic stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
	We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low, as the probability of observing 0 relevant documents in a given sample (analogous to a set of consecutive irrelevant results) is a decreasing function of the number of relevant documents in the population.
	We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; 
	2) it lacks a formal description that would help to find a suitable threshold for the criterion; and 3) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.

	Figure \ref{unseen-proportions} illustrates this third point. 
	We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
	In the top figure, machine learning (ML) has performed well, and 74\% of the screened documents were relevant. 
	In the bottom figure, ML has performed less well, and only 26\% of the screened documents were relevant.
	In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
	Recall is not simply a function of the proportion of unseen relevant documents, but also of the number of unseen documents. 
	This also means that where ML has performed well (as in the top figure), a low proportion of relevant documents in those not yet checked is indicative of lower recall than where ML has performed less well.
	Likewise, where the proportion of relevant documents in the whole corpus is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low. 
	This shows us that even a perfect estimator of the proportion of unseen relevant documents is insufficient on its own to provide sufficient information about when to stop screening. To estimate recall reliably, it is necessary to take into account the total number of unseen relevant documents (or their proportion times the number of unseen documents).

@@

	Finally, Howard et al. \cite{Howard2020} present a method to estimate recall based on the number of irrelevant documents $D$ observed in a list of documents since the $\delta th$ previous relevant document. 
	They reason that this should follow the negative binomial distribution based on the proportion of remaining relevant documents $p$, and use this information to estimate $\hat{p}$, and with this, the total number of relevant articles and the estimated recall.

	However, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours). This is not guaranteed by the criterion itself but rather a finding of the simulation with example datasets. Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$. Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
	Lastly, because screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial, even though the latter can be a good approximation for cases with large numbers of documents.

	These last examples are promising developments, but they all fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance. In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.

@@

	In our screening setup, we start off with $N_{tot}$ documents that are potentially relevant. $\rho_{tot}$ of these documents are actually relevant, but we don't know this value \textit{a priori}. As we screen relevant documents we include them, so $\rho_{seen}$ represents the number of relevant documents screened, and recall $\tau$ is given by

	\begin{equation}
	\tau = \frac{\rho_{seen}}{\rho_{tot}}
	\label{eq:recall-def}
	\end{equation}

@@

	At the start of the sample, $N_{AL}$ is the number of documents seen during the active learning process, and $N$ is the number of documents remaining, so that

	\begin{equation}
	N = N_{tot} - N_{AL}
	\end{equation}

@@

	\begin{equation}
	K = \rho_{tot} - \rho_{AL}
	\label{eq:K}
	\end{equation}

	We now take random draws from the remaining $N$ documents, and denote the number of documents drawn with $n$ and the number of relevant documents drawn with $k$. The number of relevant documents seen is updated by adding the number of relevant documents seen since sampling began to the number of relevant documents seen during active learning.

	\begin{equation}
	\rho_{seen} = \rho_{AL} + k
	\label{eq:rho_seen}
	\end{equation}

@@

	\begin{equation}
	H_0 : \tau < \tau_{tar}
	\label{eq:null_hypothesis}
	\end{equation}

@@

	\begin{equation}
	k \sim Hypergeometric(N, K, n)
	\end{equation}

	We introduce a hypothetical value for $K$, which we call $K_{tar}$. This represents the minimum number of relevant documents remaining at the start of sampling compatible with our null hypothesis that recall is below our target.

	\begin{equation}
	K_{tar} = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
	\end{equation}

	This equation is derived by combining Eqs.~\ref{eq:recall-def} and \ref{eq:rho_seen}. Because $k$ can only take integer values, $K_{tar}$ is the smallest integer that satisfies the inequality in Eq.~\ref{eq:null_hypothesis}.
	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than our hypothetical value.

@@

	We test this by calculating the probability of observing $k$ or fewer relevant documents from the hypergeometric distribution given by $K_{tar}$, using the cumulative probability mass function.

	\begin{equation}
	p = P(X \leq k)$, where $X \sim Hypergeometric(N,K_{tar},n)
	\label{eq:p-value}
	\end{equation}

	Because the cumulative probability mass function $P(X \leq k)$ is decreasing with increasing $K$, this gives the maximum probability of observing $k$ for all values of $K$ compatible with our null hypothesis. Similar arguments have been made to derive confidence intervals for estimating the parameter $K$ in the hypergeometric distribution function \citep{Buonaccorsi1987, Sahai1995} and the derivation of an equivalent criterion could use the upper limit of such a confidence interval of an estimated $K$ from the observation of $k$.

	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$. This means, we can report the likelihood that we achieve a recall above our target as being more than $\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}.

	\subsubsection*{Ranked quasi-sampling}

	We now proceed to describe a special case of the method described above which we (1) use as a heuristic in order to decide when to begin random sampling; and (2) test as an independent stopping criterion. The method works by treating batches of previously screened documents as if they were random samples.

	We calculate $p$ as above for subsets of the already screened documents. Concretely, we use subsets of documents $A_i$ by looking back to the last $i$ documents, $A_i = \{d_{N_{seen} - 1}, ..., d_{N_{seen} - i}\}$, where the documents $d$ are indexed in the order in which they have been screened. For a specific $i$, this corresponds to  random sampling beginning after seeing $i$ documents in the section above.
	Thus, we set 
	$N_{AL}$ to $i$, 
	$n$ to $N_{seen}-i$, 
	$\rho_{AL}$ to the number of relevant documents seen when $i$ documents had been seen,
	and $k$ to the number of relevant documents seen since $i$ documents had been seen, and calculate $p$ according to Eq.~\ref{eq:p-value}.
	We compute $p$ for all sets $A_i$ with $i \in {N_{seen}-1 \dots 1}$.
	This gives us a vector $\bm{p}$, representing the values of $p$ which would have been estimated at each point at which we could have stopped active learning and began random sampling. The lowest probability of our null hypothesis being true that we would have thereby obtained is given by $p_{min}$. With the vectorized implementation included in our accompanying code, these calculations are completed in less than the time it would take a human to code the next document.

	First, we use this method as a useful heuristic for deciding when to stop active learning, and switch to random sampling. For this, we choose a higher threshold for the likelihood, $p_{min} < 1-\frac{\alpha}{2}$. Second, we use the same ranked quasi-sampling as an independent stopping criterion, by continuing screening with active learning until $p_{min} < 1 - \alpha$. We present the results of this second procedure separately below.

	Given that the documents seen during active learning are ranked according to predicted relevance, they do not in fact represent a random sample. This means that the test is unlikely to be accurate. It would be reasonable to assume that the proportion of relevant documents in each ranked quasi-sample is as high if not higher than the proportion of relevant documents in the unseen documents. This assumption would make this estimator conservative. As such it works in a similar way to the criterion proposed by Howard et al. \cite{Howard2020}, although it makes use of more information and provides hypothesis testing rather than just a point estimate of recall.

@@

	\begin{figure*}
		\centering
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_hyper.pdf}
			\caption[]%
			{{\small Hypergeometric sampling \\}}    
			\label{fig:hyper}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_nrs.pdf}
			\caption[]%
			{{\footnotesize Ranked quasi-sampling}}    
			\label{fig:nrs}
		\end{subfigure}
		\vskip\baselineskip
		\begin{subfigure}[b]{0.475\textwidth}
			\centering
			\includegraphics[width=\textwidth]{2_figs_jointplot_bir.pdf}
			\caption[Network2]%Example-Image
			{{\small Baseline Sampling}}    
			\label{fig:bir}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.475\textwidth}  
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_ih_50.pdf}
			\caption[]%
			{{\small 50 consecutive irrelevant results}}    
			\label{fig:ih_50}
		\end{subfigure}
		\vskip\baselineskip
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_ih_200.pdf}
			\caption[]%
			{{\small 200 consecutive irrelevant results \\}}    
			\label{fig:ih_200}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_pf.pdf}
			\caption[]%
			{{\footnotesize \textit{a priori} knowledge}}   
			\label{fig:pf}
		\end{subfigure}

@@

	\begin{figure*}
		\centering
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_burden_hyper.pdf}
			\caption[]%
			{{\small Hypergeometric sampling \\}}    
			\label{fig:hyper_ab}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.475\textwidth}   
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_burden_nrs.pdf}
			\caption[]%
			{{\footnotesize Ranked quasi-sampling}}    
			\label{fig:nrs_ab}
		\end{subfigure}
		\hfill
		\vskip\baselineskip
		\begin{subfigure}[b]{0.475\textwidth}
			\centering
			\includegraphics[width=\textwidth]{2_figs_jointplot_burden_bir.pdf}
			\caption[Network2]%Example-Image
			{{\small Baseline Sampling}}    
			\label{fig:bir_ab}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.475\textwidth}  
			\centering 
			\includegraphics[width=\textwidth]{2_figs_jointplot_burden_ih_50.pdf}
			\caption[]%
			{{\small 50 consecutive irrelevant results}}    
			\label{fig:ih_50_ab}
		\end{subfigure}
		\vskip\baselineskip

@@

	Both the random sampling and the ranked sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. 
	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
	The ranked quasi-sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases. In theory, the ranked sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random. Based on our experiments, this assumption seems reasonable, and accounts for the higher recall. Because the ranked quasi-sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data. This accounts for the higher work savings.

@@

	\begin{figure}
		\includegraphics[width=0.9\linewidth]{2_figs_wss_nrs.pdf}
		\caption{Work saved for the ranked quasi-sampling method in each dataset. Labels show the number of relevant documents and the total number of documents. The datasets are presented in order of the number of documents. The whiskers represent the 5th and 95th percentiles. The grey line shows work savings of 5\%. }
		\label{wss}
	\end{figure}

@@

	To help explain the different work savings that were observed in our experiments, we show the distribution of work savings from our ranked quasi-sampling criterion for each dataset in figure \ref{wss}. In general, higher work savings are possible when the total number of documents is larger. However, in datasets with a low proportion of relevant documents, many documents need to be screened to achieve a high confidence that there are only few relevant documents remaining in the unseen ones. Therefore, smaller work savings are possible.

@@

	One caveat in the derivation of our criteria is that we did not address the potential problem of multiple testing formally. Such a derivation is mathematically challenging and beyond the scope of this paper. However, the performance of the criteria shows that this is of limited practical concern. Formally describing screening procedures with iterative testing should be a next step towards even more rigorous stopping criteria and should be fully worked out in future research.
